---
title: "R Notebook"
output: github_document
---

premier test de controle

```{r}
install.packages("gitcreds")
library(gitcreds)
gitcreds_set()
```


```{bash}
sudo apt-get update -y
sudo apt-get install -y libglpk-dev 
sudo apt-get install -y liblzma-dev libbz2-dev
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("BiocStyle")
BiocManager::install("Rhtslib")
```

```{r}
library("knitr")
library("BiocStyle")
.cran_packages <- c("ggplot2", "gridExtra", "devtools")
install.packages(.cran_packages) 
.bioc_packages <- c("dada2", "phyloseq", "DECIPHER", "phangorn")
BiocManager::install(.bioc_packages)
# Load packages into session, and print package version
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
```


```{r}
plot(cars)
```


```{bash}
cd ~
wget https://mothur.s3.us-east-2.amazonaws.com/wiki/miseqsopdata.zip
unzip miseqsopdata.zip
```

````{r}
set.seed(100)
miseq_path <- "./MiSeq_SOP"
list.files(miseq_path)
````

```{r}
fnFs <- sort(list.files(miseq_path, pattern="_R1_001.fastq"))
fnRs <- sort(list.files(miseq_path, pattern="_R2_001.fastq"))
```

```{r}
sampleNames <- sapply(strsplit(fnFs, "_"), `[`, 1)
fnFs <- file.path(miseq_path, fnFs)
fnRs <- file.path(miseq_path, fnRs)
fnFs[1:3]
```

## [1] "./MiSeq_SOP/F3D0_S188_L001_R1_001.fastq"   "./MiSeq_SOP/F3D1_S189_L001_R1_001.fastq"  
## [3] "./MiSeq_SOP/F3D141_S207_L001_R1_001.fastq"

```{r}
fnRs[1:3]
```
## [1] "./MiSeq_SOP/F3D0_S188_L001_R2_001.fastq"   "./MiSeq_SOP/F3D1_S189_L001_R2_001.fastq"  
## [3] "./MiSeq_SOP/F3D141_S207_L001_R2_001.fastq"

```{r}
plotQualityProfile(fnFs[1:2])
```

#Grosse diminution de la qualité

```{r}
plotQualityProfile(fnRs[1:2])
```

#Permet de définir et de garder que les séquences de bonne qualité

```{r}
filt_path <- file.path(miseq_path, "filtered")
```


```{r}
if(!file_test("-d", filt_path)) dir.create(filt_path)
filtFs <- file.path(filt_path, paste0(sampleNames, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sampleNames, "_R_filt.fastq.gz"))
```


```{r}
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(240,160),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=TRUE) 
head(out)
```
##                               reads.in reads.out
## F3D0_S188_L001_R1_001.fastq       7793      7113
## F3D1_S189_L001_R1_001.fastq       5869      5299
## F3D141_S207_L001_R1_001.fastq     5958      5463
## F3D142_S208_L001_R1_001.fastq     3183      2914
## F3D143_S209_L001_R1_001.fastq     3178      2941
## F3D144_S210_L001_R1_001.fastq     4827      4312

```{r}
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
```

# Name the derep-class objects by the sample names

```{r}
names(derepFs) <- sampleNames
names(derepRs) <- sampleNames

errF <- learnErrors(filtFs, multithread=TRUE)
```

## Initializing error rates to maximum possible estimate.
## Sample 1 - 7113 reads in 1979 unique sequences.
## Sample 2 - 5299 reads in 1639 unique sequences.
## Sample 3 - 5463 reads in 1477 unique sequences.
## Sample 4 - 2914 reads in 904 unique sequences.
## Sample 5 - 2941 reads in 939 unique sequences.
## Sample 6 - 4312 reads in 1267 unique sequences.
## Sample 7 - 6741 reads in 1756 unique sequences.
## Sample 8 - 4560 reads in 1438 unique sequences.
## Sample 9 - 15637 reads in 3590 unique sequences.
## Sample 10 - 11413 reads in 2762 unique sequences.
## Sample 11 - 12017 reads in 3021 unique sequences.
## Sample 12 - 5032 reads in 1566 unique sequences.
## Sample 13 - 18075 reads in 3707 unique sequences.
## Sample 14 - 6250 reads in 1479 unique sequences.
## Sample 15 - 4052 reads in 1195 unique sequences.
## Sample 16 - 7369 reads in 1832 unique sequences.
## Sample 17 - 4765 reads in 1183 unique sequences.
## Sample 18 - 4871 reads in 1382 unique sequences.
## Sample 19 - 6504 reads in 1709 unique sequences.
## Sample 20 - 4314 reads in 897 unique sequences.
##    selfConsist step 2 
##    selfConsist step 3 
##    selfConsist step 4 
##    selfConsist step 5 
## 
## 
## Convergence after  5  rounds.
## Total reads used:  139642

```{r}
errR <- learnErrors(filtRs, multithread=TRUE)
```

## Initializing error rates to maximum possible estimate.
## Sample 1 - 7113 reads in 1660 unique sequences.
## Sample 2 - 5299 reads in 1349 unique sequences.
## Sample 3 - 5463 reads in 1335 unique sequences.
## Sample 4 - 2914 reads in 853 unique sequences.
## Sample 5 - 2941 reads in 880 unique sequences.
## Sample 6 - 4312 reads in 1286 unique sequences.
## Sample 7 - 6741 reads in 1803 unique sequences.
## Sample 8 - 4560 reads in 1265 unique sequences.
## Sample 9 - 15637 reads in 3414 unique sequences.
## Sample 10 - 11413 reads in 2522 unique sequences.
## Sample 11 - 12017 reads in 2771 unique sequences.
## Sample 12 - 5032 reads in 1415 unique sequences.
## Sample 13 - 18075 reads in 3290 unique sequences.
## Sample 14 - 6250 reads in 1390 unique sequences.
## Sample 15 - 4052 reads in 1134 unique sequences.
## Sample 16 - 7369 reads in 1635 unique sequences.
## Sample 17 - 4765 reads in 1084 unique sequences.
## Sample 18 - 4871 reads in 1161 unique sequences.
## Sample 19 - 6504 reads in 1502 unique sequences.
## Sample 20 - 4314 reads in 732 unique sequences.
##    selfConsist step 2 
##    selfConsist step 3 
##    selfConsist step 4 
##    selfConsist step 5 
##    selfConsist step 6 
## 
## 
## Convergence after  6  rounds.
## Total reads used:  139642

```{r}
plotErrors(errF)
plotErrors(errR)
```

```{r}
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
```
## Sample 1 - 7113 reads in 1979 unique sequences.
## Sample 2 - 5299 reads in 1639 unique sequences.
## Sample 3 - 5463 reads in 1477 unique sequences.
## Sample 4 - 2914 reads in 904 unique sequences.
## Sample 5 - 2941 reads in 939 unique sequences.
## Sample 6 - 4312 reads in 1267 unique sequences.
## Sample 7 - 6741 reads in 1756 unique sequences.
## Sample 8 - 4560 reads in 1438 unique sequences.
## Sample 9 - 15637 reads in 3590 unique sequences.
## Sample 10 - 11413 reads in 2762 unique sequences.
## Sample 11 - 12017 reads in 3021 unique sequences.
## Sample 12 - 5032 reads in 1566 unique sequences.
## Sample 13 - 18075 reads in 3707 unique sequences.
## Sample 14 - 6250 reads in 1479 unique sequences.
## Sample 15 - 4052 reads in 1195 unique sequences.
## Sample 16 - 7369 reads in 1832 unique sequences.
## Sample 17 - 4765 reads in 1183 unique sequences.
## Sample 18 - 4871 reads in 1382 unique sequences.
## Sample 19 - 6504 reads in 1709 unique sequences.
## Sample 20 - 4314 reads in 897 unique sequences.
```{r}
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
```

## Sample 1 - 7113 reads in 1660 unique sequences.
## Sample 2 - 5299 reads in 1349 unique sequences.
## Sample 3 - 5463 reads in 1335 unique sequences.
## Sample 4 - 2914 reads in 853 unique sequences.
## Sample 5 - 2941 reads in 880 unique sequences.
## Sample 6 - 4312 reads in 1286 unique sequences.
## Sample 7 - 6741 reads in 1803 unique sequences.
## Sample 8 - 4560 reads in 1265 unique sequences.
## Sample 9 - 15637 reads in 3414 unique sequences.
## Sample 10 - 11413 reads in 2522 unique sequences.
## Sample 11 - 12017 reads in 2771 unique sequences.
## Sample 12 - 5032 reads in 1415 unique sequences.
## Sample 13 - 18075 reads in 3290 unique sequences.
## Sample 14 - 6250 reads in 1390 unique sequences.
## Sample 15 - 4052 reads in 1134 unique sequences.
## Sample 16 - 7369 reads in 1635 unique sequences.
## Sample 17 - 4765 reads in 1084 unique sequences.
## Sample 18 - 4871 reads in 1161 unique sequences.
## Sample 19 - 6504 reads in 1502 unique sequences.
## Sample 20 - 4314 reads in 732 unique sequences.

```{r}
dadaFs[[1]]
```

## dada-class: object describing DADA2 denoising results
## 128 sample sequences were inferred from 1979 input unique sequences.
## Key parameters: OMEGA_A = 1e-40, BAND_SIZE = 16, USE_QUALS = TRUE

```{r}
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs)

seqtabAll <- makeSequenceTable(mergers[!grepl("Mock", names(mergers))])
table(nchar(getSequences(seqtabAll)))
```
## 
## 251 252 253 254 255 
##   1  84 182   5   2

```{r}
seqtabNoC <- removeBimeraDenovo(seqtabAll)
```

#Assignement taxonomique

```{r}
fastaRef <- "./rdp_train_set_16.fa.gz"
taxTab <- assignTaxonomy(seqtabNoC, refFasta = fastaRef, multithread=TRUE)
unname(head(taxTab))
```


```{bash}
cd
wget  https://zenodo.org/record/4587955/files/silva_nr99_v138.1_train_set.fa.gz
```

```{r}
fastaRef <-"/home/rstudio/silva_nr99_v138.1_train_set.fa.gz"
taxTab<-assignTaxonomy(seqtabNoC, refFasta=fastaRef, multithread=TRUE)
unname(head(taxTab))
```

```{r}
seqs <- getSequences(seqtabNoC)
names(seqs) <- seqs # This propagates to the tip labels of the tree
alignment <- AlignSeqs(DNAStringSet(seqs), anchor=NA,verbose=FALSE)
```

```{r}
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
```

```{r}
phangAlign <- phyDat(as(alignment, "matrix"), type="DNA")
dm <- dist.ml(phangAlign)
treeNJ <- NJ(dm) # Note, tip order != sequence order
fit = pml(treeNJ, data=phangAlign)
fitGTR <- update(fit, k=4, inv=0.2)
fitGTR <- optim.pml(fitGTR, model="GTR", optInv=TRUE, optGamma=TRUE,
        rearrangement = "stochastic", control = pml.control(trace = 0))
detach("package:phangorn", unload=TRUE)
```

```{r}
samdf <- read.csv("https://raw.githubusercontent.com/spholmes/F1000_workflow/master/data/MIMARKS_Data_combined.csv",header=TRUE)
samdf$SampleID <- paste0(gsub("00", "", samdf$host_subject_id), "D", samdf$age-21)
samdf <- samdf[!duplicated(samdf$SampleID),] # Remove dupicate entries for reverse reads
rownames(seqtabAll) <- gsub("124", "125", rownames(seqtabAll)) # Fix discrepancy
all(rownames(seqtabAll) %in% samdf$SampleID) # TRUE
```


```{r}
rownames(samdf) <- samdf$SampleID
keep.cols <- c("collection_date", "biome", "target_gene", "target_subfragment",
"host_common_name", "host_subject_id", "age", "sex", "body_product", "tot_mass",
"diet", "family_relationship", "genotype", "SampleID") 
samdf <- samdf[rownames(seqtabAll), keep.cols]
```

```{r}
ps <- phyloseq(otu_table(seqtabNoC, taxa_are_rows=FALSE), 
               sample_data(samdf), 
               tax_table(taxTab),phy_tree(fitGTR$tree))
ps <- prune_samples(sample_names(ps) != "Mock", ps) # Remove mock sample
ps
```

#loading les données

```{r}
import_biom
import_qiime

ps_connect <-url("https://raw.githubusercontent.com/spholmes/F1000_workflow/master/data/ps.rds")
ps = readRDS(ps_connect)
ps
```

```{r}
rank_names(ps)

table(tax_table(ps)[, "Phylum"], exclude = NULL)
ps <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized"))

prevdf = apply(X = otu_table(ps),
               MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
               
prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(ps),
                    tax_table(ps))

plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),sum(df1$Prevalence))})
```

```{r}
filterPhyla = c("Fusobacteria", "Deinococcus-Thermus")

ps1 = subset_taxa(ps, !Phylum %in% filterPhyla)
ps1
```

```{r}
prevdf1 = subset(prevdf, Phylum %in% get_taxa_unique(ps1, "Phylum"))
ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(ps),color=Phylum)) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +  geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")
```
# Define prevalence threshold as 5% of total samples


```{r}
prevalenceThreshold = 0.05 * nsamples(ps)
prevalenceThreshold
```

```{r}
install.packages("gitcreds")
library(gitcreds)
gitcreds_set()
```

```{r}
keepTaxa = rownames(prevdf1)[(prevdf1$Prevalence >= prevalenceThreshold)]
ps2 = prune_taxa(keepTaxa, ps)
```

```{r}
length(get_taxa_unique(ps2, taxonomic.rank = "Genus"))
ps3 = tax_glom(ps2, "Genus", NArm = TRUE)
h1 = 0.4
ps4 = tip_glom(ps2, h = h1)
```

```{r}
multiPlotTitleTextSize = 15
p2tree = plot_tree(ps2, method = "treeonly",
                   ladderize = "left",
                   title = "Before Agglomeration") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))
p3tree = plot_tree(ps3, method = "treeonly",
                   ladderize = "left", title = "By Genus") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))
p4tree = plot_tree(ps4, method = "treeonly",
                   ladderize = "left", title = "By Height") +
  theme(plot.title = element_text(size = multiPlotTitleTextSize))
```

```{r}
grid.arrange(nrow = 1, p2tree, p3tree, p4tree)
```
```{r}
plot_abundance = function(physeq,title = "",
                          Facet = "Order", Color = "Phylum"){
  p1f = subset_taxa(physeq, Phylum %in% c("Firmicutes"))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "sex",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}
```

```{r}
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})
```

```{r}
plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
grid.arrange(nrow = 2,  plotBefore, plotAfter)
```
```{r}
psOrd = subset_taxa(ps3ra, Order == "Lactobacillales")
plot_abundance(psOrd, Facet = "Genus", Color = NULL)
```

```{r}
.cran_packages <- c( "shiny","miniUI", "caret", "pls", "e1071", "ggplot2", "randomForest", "dplyr", "ggrepel", "nlme", "devtools",
                  "reshape2", "PMA", "structSSI", "ade4",
                  "ggnetwork", "intergraph", "scales")
.github_packages <- c("jfukuyama/phyloseqGraphTest")
.bioc_packages <- c("genefilter", "impute")
```
```{r}
.inst <- .cran_packages %in% installed.packages()
if (any(!.inst)){
  install.packages(.cran_packages[!.inst],repos = "http://cran.rstudio.com/")
}
```
```{r}
.inst <- .github_packages %in% installed.packages()
if (any(!.inst)){
  devtools::install_github(.github_packages[!.inst])
}
```

```{r}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)){BiocManager::install(.bioc_packages[!.inst])
}
```


```{r}
qplot(sample_data(ps)$age, geom = "histogram",binwidth=20) + xlab("age")
```
```{r}
qplot(log10(rowSums(otu_table(ps))),binwidth=0.2) +
  xlab("Logged counts-per-sample")
```

```{r}
sample_data(ps)$age_binned <- cut(sample_data(ps)$age,
                          breaks = c(0, 100, 200, 400))
levels(sample_data(ps)$age_binned) <- list(Young100="(0,100]", Mid100to200="(100,200]", Old200="(200,400]")
sample_data(ps)$family_relationship=gsub(" ","",sample_data(ps)$family_relationship)
pslog <- transform_sample_counts(ps, function(x) log(1 + x))
out.wuf.log <- ordinate(pslog, method = "MDS", distance = "wunifrac")
evals <- out.wuf.log$values$Eigenvalues
plot_ordination(pslog, out.wuf.log, color = "age_binned") +
  labs(col = "Binned Age") +
  coord_fixed(sqrt(evals[2] / evals[1]))
```
```{r}
rel_abund <- t(apply(otu_table(ps), 1, function(x) x / sum(x)))
qplot(rel_abund[, 12], geom = "histogram",binwidth=0.05) +
  xlab("Relative abundance")
```
#ordination


```{r}
outliers <- c("F5D165", "F6D165", "M3D175", "M4D175", "M5D175", "M6D175")
ps <- prune_samples(!(sample_names(ps) %in% outliers), ps)
which(!rowSums(otu_table(ps)) > 1000)
```
```{r}
ps <- prune_samples(rowSums(otu_table(ps)) > 1000, ps)
pslog <- transform_sample_counts(ps, function(x) log(1 + x))
```

```{r}
out.pcoa.log <- ordinate(pslog,  method = "MDS", distance = "bray")
evals <- out.pcoa.log$values[,1]
plot_ordination(pslog, out.pcoa.log, color = "age_binned",
                  shape = "family_relationship") +
  labs(col = "Binned Age", shape = "Litter")+
  coord_fixed(sqrt(evals[2] / evals[1]))
```

```{r}
out.dpcoa.log <- ordinate(pslog, method = "DPCoA")
evals <- out.dpcoa.log$eig
plot_ordination(pslog, out.dpcoa.log, color = "age_binned", label= "SampleID",
                  shape = "family_relationship") +
  labs(col = "Binned Age", shape = "Litter")+
  coord_fixed(sqrt(evals[2] / evals[1]))
```
```{r}
plot_ordination(pslog, out.dpcoa.log, type = "species", color = "Phylum") +
  coord_fixed(sqrt(evals[2] / evals[1]))
```
```{r}
out.wuf.log <- ordinate(pslog, method = "PCoA", distance ="wunifrac")
evals <- out.wuf.log$values$Eigenvalues
plot_ordination(pslog, out.wuf.log, color = "age_binned",
                  shape = "family_relationship") +
  coord_fixed(sqrt(evals[2] / evals[1])) +
  labs(col = "Binned Age", shape = "Litter")
```
```{r}
abund <- otu_table(pslog)
abund_ranks <- t(apply(abund, 1, rank))
```

```{r}
abund_ranks <- abund_ranks - 329
abund_ranks[abund_ranks < 1] <- 1
```

```{r}
library(dplyr)
library(reshape2)
abund_df <- melt(abund, value.name = "abund") %>%
  left_join(melt(abund_ranks, value.name = "rank"))
colnames(abund_df) <- c("sample", "seq", "abund", "rank")

abund_df <- melt(abund, value.name = "abund") %>%
  left_join(melt(abund_ranks, value.name = "rank"))
colnames(abund_df) <- c("sample", "seq", "abund", "rank")

sample_ix <- sample(1:nrow(abund_df), 8)
ggplot(abund_df %>%
         filter(sample %in% abund_df$sample[sample_ix])) +
  geom_point(aes(x = abund, y = rank, col = sample),
             position = position_jitter(width = 0.2), size = 1.5) +
  labs(x = "Abundance", y = "Thresholded rank") +
  scale_color_brewer(palette = "Set2")
```
```{r}
library(ade4)
ranks_pca <- dudi.pca(abund_ranks, scannf = F, nf = 3)
row_scores <- data.frame(li = ranks_pca$li,
                         SampleID = rownames(abund_ranks))
col_scores <- data.frame(co = ranks_pca$co,
                         seq = colnames(abund_ranks))
tax <- tax_table(ps) %>%
  data.frame(stringsAsFactors = FALSE)
tax$seq <- rownames(tax)
main_orders <- c("Clostridiales", "Bacteroidales", "Lactobacillales",
                 "Coriobacteriales")
tax$Order[!(tax$Order %in% main_orders)] <- "Other"
tax$Order <- factor(tax$Order, levels = c(main_orders, "Other"))
tax$otu_id <- seq_len(ncol(otu_table(ps)))
row_scores <- row_scores %>%
  left_join(sample_data(pslog))
col_scores <- col_scores %>%
  left_join(tax)
```
```{r}
evals_prop <- 100 * (ranks_pca$eig / sum(ranks_pca$eig))
ggplot() +
  geom_point(data = row_scores, aes(x = li.Axis1, y = li.Axis2), shape = 2) +
  geom_point(data = col_scores, aes(x = 25 * co.Comp1, y = 25 * co.Comp2, col = Order),
             size = .3, alpha = 0.6) +
  scale_color_brewer(palette = "Set2") +
  facet_grid(~ age_binned) +
  guides(col = guide_legend(override.aes = list(size = 3))) +
  labs(x = sprintf("Axis1 [%s%% variance]", round(evals_prop[1], 2)),
       y = sprintf("Axis2 [%s%% variance]", round(evals_prop[2], 2))) +
  coord_fixed(sqrt(ranks_pca$eig[2] / ranks_pca$eig[1])) +
  theme(panel.border = element_rect(color = "#787878", fill = alpha("white", 0)))
```
```{r}
ps_ccpna <- ordinate(pslog, "CCA", formula = pslog ~ age_binned + family_relationship)
```
```{r}
library(ggrepel)
ps_scores <- vegan::scores(ps_ccpna)
sites <- data.frame(ps_scores$sites)
sites$SampleID <- rownames(sites)
sites <- sites %>%
  left_join(sample_data(ps))

species <- data.frame(ps_scores$species)
species$otu_id <- seq_along(colnames(otu_table(ps)))
species <- species %>%
  left_join(tax)
evals_prop <- 100 * ps_ccpna$CCA$eig[1:2] / sum(ps_ccpna$CA$eig)
ggplot() +
  geom_point(data = sites, aes(x = CCA1, y = CCA2), shape = 2, alpha = 0.5) +
  geom_point(data = species, aes(x = CCA1, y = CCA2, col = Order), size = 0.5) +
  geom_text_repel(data = species %>% filter(CCA2 < -2),
                    aes(x = CCA1, y = CCA2, label = otu_id),
            size = 1.5, segment.size = 0.1) +
  facet_grid(. ~ family_relationship) +
  guides(col = guide_legend(override.aes = list(size = 3))) +
  labs(x = sprintf("Axis1 [%s%% variance]", round(evals_prop[1], 2)),
        y = sprintf("Axis2 [%s%% variance]", round(evals_prop[2], 2))) +
  scale_color_brewer(palette = "Set2") +
  coord_fixed(sqrt(ps_ccpna$CCA$eig[2] / ps_ccpna$CCA$eig[1])*0.45   ) +
  theme(panel.border = element_rect(color = "#787878", fill = alpha("white", 0)))
```



